--- 
title: "R-Course2"
author: "Rob Hayward"
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
output: bookdown::gitbook
documentclass: book
bibliography: [book.bib, packages.bib]
biblio-style: apalike
link-citations: yes
github-repo: rstudio/bookdown-demo
description: "This is a minimal example of using the bookdown package to write a book. The output format for this example is bookdown::gitbook."
---

# Index


<!--chapter:end:index.Rmd-->

# Introduction {#intro}

R is an open-source statistical programming language.  It is available from the following website: [R-Cran](http://www.r-project.org/). You should install R on you laptop.  For us it is probably best to use the [following link](http://www.stats.bris.ac.uk/R/).  You can also install [R-Studio](https://www.rstudio.com/) which is an Integrated Development Environment (IDE) for R. 

##Competitors

R is one of the main programmable statistical languages.  Others include:
* Stata:  Use in academic econometrics. This is not open source.
* Matlab:  Used extensively in academic finance.  It is not open-source.  There is an open-source clone called *Octave*.  It can be downloaded [here.](https://www.gnu.org/software/octave/)
* Python: Computer Language for statistical programming.  
* Julia:  New language that is faster than R
* C++:  Programming language.  Underneath R is C++ in some cases. 

##Advantages of using R
The main advantages of using an open-source statistical programming languages like R are:
* It can be used by anyone for no cost.  Your model will not depend on the next institution having the same software licence
* Programming language will encourage *reproducible research*

Reproducible Research means that results can be easily replicated.  This means
* Mistakes are less likely to happen than would be the case with *copy and paste*
* Mistakes can be more swiftly identified
* Changes to data or research methods are easier to make

##Additional Information
Some additional information on R, R-Studio and Reproducible Research 
* [LSE commentary on Reinhart and Rogoff](http://blogs.lse.ac.uk/impactofsocialsciences/2013/10/23/open-data-in-economics-the-basis-of-reproducible-research/)

<!--chapter:end:01-Intro.Rmd-->

# First R

##Introduction
All the regular mathematical operations can be carried out. 
```{r, eval=FALSE}

4 + 5
4 * 5
4 / 5
4^5

sqrt(5)
log(5)

test <- 2*3
test

test = 2/3
test
```

**Practice**

* Calculate the average of 2, 5, 6, 8, 9, 15, 1, 3
* Calculate 10405 10222 as a percentage change


## Objects
R uses objects. The main objects are:

* numeric
* character
* logical 
* vector
* data frame


When things go wrong, it is often the case that you are to something impossible with that object.  For example, adding up characters. 


###Character
For an example of character
```{r}
a <- "My name"
a
```

**Practice**

* Assign you first and last name to the objects ``a`` and ``b``
* Add ``a`` to ``b``

###Logical (boolean,  true or false)
```{r}
1 == 1

1 == 2 

1 & 2 == 1

1 | 2 == 1
```

**Practice**

* Test ``Myname`` is equal to "Rob Hayward"


## Vectors and matrices
Strings or numerics can be concatenated or combined to form a vector or matrix
```{r}  
mynumbers <- c(3,5,6,7,9)
mynumbers
mynumbers <- c(1:10)
mynumbers
```
To create a sequence, use the command sequence
```{r}
a <- seq(0.5,2.5, length=100)
a
b <- seq(1,10,1)
b
mynumbers <- 1:12
m <-matrix(mynumbers, nrow=4)
m
```

Individual components of a vector, matrix or data.frame can be identifies by using square brackets and a pair of numbers with the first equal to the row and the second equal to column.  You only need the first number for a vector. 

```{r}
A <- c(1, 2, 6, 8, 10)
B <- c(55, 56, 60, 20, 18)
C <- data.frame(A, B)
C[3, 2]
C$A
```

As you can see, a ``$`` can also be used to identify a column by `name'. 

**Practice**

* Assign a two-by-two matrix with the numbers from 1 to 4 to the letter ``c``
* Make ``Mynumber`` equal to the bottom left corner
* Make ``Mysecondnumber`` equal to the top right corner
* Assign a two-by-two matrix with the numbers from 5 to 8 to the letter ``d``
* Multiply ``c`` and ``d``


Matrices and vectors can only contain one object. Try the previous exercise with the ``d`` matrix changes to the sequence ``a, b, c, d``. 

Data-frames can group together vectors or various class.  These will probably be the most useful for us as we want time series with dates and integers. 


##Plotting
The plot function will allow you to create graphs with the data that you have. 
```{r, fig.height=4}
x<-seq(-pi,pi,length=10)
y<-sin(x)
plot(x,y)
lines(x,y)
```
If you change the length of the x variable to 100 you will get a smoother line. 
```{r, fig.height=4}
x<-seq(-pi,pi,length=100)
y<-sin(x)
plot(x,y)
lines(x,y)
```
```{r, fig.height=4}
z<-rnorm(100)
hist(z, prob=TRUE)
mu<-mean(z)
sig<-sd(z)
x<-seq(-4,4,length=500)
y<-dnorm(x,mu,sig)
lines(x,y, col='red')
```

##Input data
By far the easiest way to input data is to read the data from a csv file.  There are packages for reading from excel and from other databases that we will look at later. 

```{r}
da <- read.csv('./Data/bac.csv')
head(da)
str(da)
```
It is important to make sure that the variables in the data.frame of of the correct class.  In this case, the ``Date`` series (first column) is a ``factor``.  We need to change it to date. 

```{r}
da$Date <- as.Date(da$Date, format = "%d/%m/%Y")
str(da)
```

**Practice**

* Go to Yahoo finance and download data for Ford (F) from January 2000 to December 2017
* Save the csv file to the desktop and import into R
* Plot the Closing price
* Beautify the plot.


<!--chapter:end:02-FirstR.Rmd-->

--- 
title: "R Course"
author: "Rob Hayward"
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
output: bookdown::gitbook
documentclass: book
bibliography: [book.bib, packages.bib]
biblio-style: apalike
link-citations: yes
github-repo: robhayward/R-Course2
description: "This is a practice at the R Course"
---

# Prerequisites

Placeholder



<!--chapter:end:index.Rmd-->

# Introduction {#intro}

========================================================
R is an open-source statistical programming language.  It is available from the following website: [R-Cran](http://www.r-project.org/). You should install R on you laptop.  For us it is probably best to use the [following link](http://www.stats.bris.ac.uk/R/).  You can also install [R-Studio](https://www.rstudio.com/) which is an Integrated Development Environment (IDE) for R. 

Competitors

----------------------------------------------------------
R is one of the main programmable statistical languages.  Others include:
* Stata:  Use in academic econometrics. This is not open source.
* Matlab:  Used extensively in academic finance.  It is not open-source.  There is an open-source clone called *Octave*.  It can be downloaded [here.](https://www.gnu.org/software/octave/)
* Python: Computer Language for statistical programming.  
* Julia:  New language that is faster than R
* C++:  Programming language.  Underneath R is C++ in some cases. 

Advantages of using R
------------------------------------------------------------
The main advantages of using an open-source statistical programming languages like R are:
* It can be used by anyone for no cost.  Your model will not depend on the next institution having the same software licence
* Programming language will encourage *reproducible research*

Reproducible Research
--------------------------------------------------------------
Reproducible Research means that results can be easily replicated.  This means
* Mistakes are less likely to happen than would be the case with *copy and paste*
* Mistakes can be more swiftly identified
* Changes to data or research methods are easier to make

Additional Information
---------------------------------------------------------
Some additional information on R, R-Studio and Reproducible Research 
* [LSE commentary on Reinhart and Rogoff](http://blogs.lse.ac.uk/impactofsocialsciences/2013/10/23/open-data-in-economics-the-basis-of-reproducible-research/)

<!--chapter:end:01-intro.Rmd-->


# First R

Placeholder



<!--chapter:end:02-literature.Rmd-->


<!--chapter:end:R-Course2.Rmd-->

